---
title: "Final Project"
author: "Seyedehbahareh Hashemimovahed"
date: "2023-05-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(plyr)
library(matrixStats)
library(plotly)
```



### NFL Divisional and None-Divisional Game Analysis Project
#### We are comparing the Divisional and Non-devisional games in NFL and finding out if divisional games have a different total final scores and points. 


##### First, we will be reading the datasets provided to us 

```{r}
#Read the data
nfl_scores <- read.csv('spreadspoke_scores.csv')
nfl_scores <- drop_na(nfl_scores)
nfl_scores <- filter(nfl_scores, schedule_season >= 2002)

nfl_teams <- read.csv('nfl_teams.csv')
nfl_teams <- select(nfl_teams, team_name, team_division)
```

##### Now we will be Merging the tables scores and team information to get the division of each team for each game.

```{r}
nfl_data <- merge(x = nfl_scores, y = nfl_teams, by.x = 'team_home', by.y = 'team_name', all.x = TRUE)
nfl_data <- rename(nfl_data, team_division_home = team_division)
nfl_data <- merge(x = nfl_data, y = nfl_teams, by.x = 'team_away', by.y = 'team_name', all.x = TRUE)
nfl_data <- rename(nfl_data, team_division_away = team_division)
```

##### We need to fix the data by filling in the missing values for Cities.

```{r}
A <- filter(nfl_data, team_away == 'Las Vegas Raiders')
A$team_division_away <- 'AFC West'
B <- filter(nfl_data, team_away == 'St. Louis Rams')
B$team_division_away <- 'NFC West'
C <- filter(nfl_data, team_home == 'St. Louis Rams')
nfl_data <- filter(nfl_data, team_away != 'Las Vegas Raiders')
nfl_data <- filter(nfl_data, team_away != 'St. Louis Rams')
nfl_data <- filter(nfl_data, team_home != 'St. Louis Rams')
nfl_data <- rbind(nfl_data, A, B, C)
```

##### Here, we are finding the point difference between teams at end of each game and get total for divisional games. 

```{r}
nfl_data <- mutate(nfl_data, pd = abs(score_home-score_away))
nfl_data <- mutate(nfl_data, total = score_home + score_away)
nfl_data <- mutate(nfl_data, divisional_game = ifelse(team_division_home == team_division_away, TRUE, FALSE))

head(nfl_data,10)
```

###Question 1 
####Are point differentials different for divisional games vs non-divisional?
###Expectation
####point differentials usually don't differ in divisonal games vs non-divisonal games

```{r}
df1 <- nfl_data %>%
  group_by(divisional_game) %>%
  summarize(mean_point_diff = mean(abs(score_home - score_away)))
summary(df1)

df1 <- lm(pd~divisional_game, data = nfl_data)

summary(df1)

# Create a box plot of point differentials by game type

ggplot(nfl_data, aes(x = divisional_game, y = abs(score_home - score_away))) +
  geom_boxplot() +
  labs(x = "Divisional Game", y = "Point Differential")

ggplot(nfl_data, aes(x = divisional_game, y = abs(score_home - score_away))) +
  geom_point() +
  labs(x = "Divisional Game", y = "Point Differential")
```

###Question 2
####Are spreads diferent for divisional games?
###Expectation
####divisional games have a lower spreads.
```{r}
df2 <- nfl_data %>%
  group_by(divisional_game) %>%
  summarize(mean_spread = mean(abs(score_home - score_away) - abs(spread_favorite)))
summary(df2)


df2 <- lm(spread_favorite~divisional_game, data = nfl_data)
summary(df2)

ggplot(nfl_data, aes(x = divisional_game, y = abs(score_home - score_away) - abs(spread_favorite))) +
  geom_boxplot() +
  labs(x = "Divisional Game", y = "Spread Differential")


```


###Question 3
####Are game totals different for divisional games?
###Expectation
####yes, divisional games have 1.6 points diffrence and lower totals than non-divisional games. 
```{r}

df3 <- nfl_data %>%
  group_by(divisional_game) %>%
  summarize(mean_game_total = mean(score_home + score_away))
summary(df3)

df3 <- lm(total~divisional_game, data = nfl_data)
summary(df3)

ggplot(nfl_data, aes(x = divisional_game, y = score_home + score_away)) +
  geom_boxplot() +
  labs(x = "Divisional Game", y = "Game Total")

```


###Question 4
####Are over/under lines diferent for divisional games
###Expectation
#####divisional games have a lower over_unders compared to none_divisional games.
```{r}
df4 <- nfl_data %>%
  group_by(divisional_game) %>%
  summarize(mean_over_under = mean(over_under_line))

summary(df4)
df4 <- lm(over_under_line~divisional_game, data = nfl_data)
summary(df4)
```
#### Conclusion: 
##### It seems that the point spread between the divisional and none-devisional games for NFL games are lower for Divisional matchups.   
